
\resetcounters

\bibliographystyle{asp2010}

\markboth{Anderson, Rosolowsky, and Dowler}{CyberSKA}


\title{\ssindex{projects!CyberSKA}CyberSKA \ssindex{astronomy!radio}Radio Imaging Metadata and VO Compliance Engineering}
\author{K.~R.~Anderson,$^1$ E.~Rosolowsky,$^1$ and P.~Dowler$^2$
\affil{$^1$University of British Columbia, Okanagan, 3333 University Way
Kelowna, BC}
\affil{$^2$National Research Council Canada, 1200 Montreal Road
Building M-58, Ottawa, Ontario}}

\aindex{Anderson, K. R.}
\aindex{Rosolowsky, E.}
\aindex{Dowler, P.}

\begin{abstract}
The \ssindex{projects!CyberSKA}CyberSKA project has written a specification for the \ssindex{data!metadata}metadata encapsulation of \ssindex{astronomy!radio}radio astronomy data products pursuant to insertion into the\ssindex{Virtual Observatory(VO)} VO-compliant Common Archive Observation Model (\ssindex{data!model!CAOM}CAOM) database hosted by the Canadian Astronomy Data Centre (\ssindex{data centers!Canadian Astronomy Data Centre (CADC)}CADC). This specification accommodates \ssindex{astronomy!radio}radio \ssindex{data formats!FITS}FITS Image and UV Visibility data, as well as pure \ssindex{packages!CASA}CASA\ooindex{CASA, ascl:1107.013} Tables Imaging and Visibility Measurement Sets. To extract and engineer radio \ssindex{data!metadata}metadata, we have authored two software packages: metaData (v0.5.0) and mddb (v1.3). Together, these \ssindex{computer languages!Python}python packages can convert all the above stated data format types into concise \ssindex{data formats!FITS}FITS-like header files, engineer the \ssindex{data!metadata}metadata to conform to the \ssindex{data!model!CAOM}CAOM data model, and then insert these engineered data into the \ssindex{data centers!Canadian Astronomy Data Centre (CADC)}CADC database, which subsequently becomes published through the Canadian \ssindex{Virtual Observatory (VO)}Virtual Observatory. The metaData and mddb packages have, for the first time, published \ssindex{observatories!Earth-based!ALMA}ALMA imaging data on\ssindex{Virtual Observatory(VO)} VO services. Our ongoing work aims to integrate visibility data from \ssindex{observatories!Earth-based!ALMA}ALMA and the \ssindex{observatories!Earth-based!SKA}SKA into\ssindex{Virtual Observatory(VO)} VO services and to enable user-submitted \ssindex{astronomy!radio}radio data to move seamlessly into the \ssindex{Virtual Observatory (VO)}Virtual Observatory. 
\end{abstract}

\section{Introduction}
The \ssindex{projects!CyberSKA}CyberSKA project  (\url{http://cyberska.org/}) is developing a Data Query Service (DQS) for users of the web portal. It was specified that this service should be compliant with \ssindex{Virtual Observatory (VO)}Virtual Observatory (VO) protocols in order that \ssindex{projects!CyberSKA}CyberSKA datasets are exposed through\ssindex{Virtual Observatory(VO)} VO services. In a \textit{pro tem} ``proof of concept'' effort, \ssindex{projects!CyberSKA}CyberSKA project collaborated with \ssindex{data centers!Canadian Astronomy Data Centre (CADC)}CADC in order to leverage extant\ssindex{Virtual Observatory(VO)} VO services already available at the Centre.

To this end, the authors have developed  specifications and software that both encapsulate the necessary \ssindex{data!metadata}metadata associated with datasets uploaded to the  \ssindex{projects!CyberSKA}CyberSKA Data Management Service (DMS), and then engineer that \ssindex{data!metadata}metadata to conform to the \ssindex{data!model!CAOM}CAOM data model \citep{dowler_2007}. The engineered data are then inserted into the \ssindex{data!model!CAOM}CAOM database. Once in the \ssindex{data!model!CAOM}CAOM database, the data then become available through various\ssindex{Virtual Observatory(VO)} VO services, such as the Data Discovery Tool. 

\section{The Software}
Pursuant to the goals outlined above, two software packages have been developed to date: \ssindex{data!metadata}metaData, v0.5.0 and mddb, v1.3. As the name suggests, the metaData package is responsible for encapsulating the \ssindex{data!metadata}metadata of a variety of data formats and types.\footnote{These are data ``types'' in a meta sense, such as ``imaging'' and ``visibility.''} \ssindex{data!metadata}Metadata are extracted and encapsulated only and explicitly for \ssindex{data formats!FITS}FITS images, UVFITS data, \ssindex{packages!CASA}Casa\ooindex{CASA, ascl:1107.013} Tables images, and Casa Tables Visibility Measurement Sets. The mddb package is then responsible for configuring thusly encapsulated \ssindex{data!metadata}metadata for \ssindex{data!model!CAOM}CAOM database insertion.

\subsection{metaData, v0.5.0}
The metaData package \citep{and_2011} is arranged in an hierarchy of ``main'' modules that provide the classes for determining the ``MIME type'' of the previously mentioned data types. Though not strictly MIME, these types have been specified by the \ssindex{projects!CyberSKA}CyberSKA group in order to explicitly identify a dataset type. The metaData mime-typing facility can identify the previously mentioned data formats and types,\footnote{Future implementations of metaData MimeTyping and Handler classes could potentially accommodate future adopted formats such as \ssindex{data formats!HDF}HDF5.} and tag a \ssindex{projects!CyberSKA}CyberSKA dataset as one of the following:
\begin{itemize}
\item ``image/fits-image''
\item ``image/fits-uvw''
\item ``image/ms-image''
\item ``image/ms-uvw''
\end{itemize}
Current metaData MimeTyping and Handlers classes are outlined in Table \ref{tab:mods}.
\begin{table}[htbp]
  \centering
  \begin{tabular}{|l|p{9.0cm}|} 
    \hline
    \sc \textbf{Module} & \textbf{Classes} \\
    \hline
    fitsMimeTyping & FITSMimeTyping, FITSMimeTypeError(TypeError) \\
    msMimeTyping   & MSMimeTyping,   MimetypeError(TypeError) \\
    msHandlers     & MSHandlers,     MSTableValueError(AttributeError) \\
    fitsHandlers   & FitsHandlers \\
    casaImageHandlers  & CasaImageHandlers, CasaImageTypeError(TypeError)\\
    \hline
  \end{tabular}
  \caption{metaData Module and Class relationship}
  \label{tab:mods}
\end{table}

A set of support modules provide these classes with generalised functional utility, shown in Table \ref{tab:subdirs}.
\begin{table}[htbp]
  \centering
  \begin{tabular}{|l|p{9.0cm}|} 
    \hline
    \sc \textbf{Support Directory} & \textbf{Modules} \\
    \hline
    convert/ & polarizationConversions, mjdConversions, timeConversions,
               directionConversions, frequencyConversions \\
    incl/   &  imageInclusion, tablesInclusion \\
    utils/  &  genUtils, runUtils \\
    \hline
    docs/   &  *.html\\
    \hline
  \end{tabular}
  \caption{metaData subdirectories and support modules}
  \label{tab:subdirs}
\end{table}

The modules found in the convert/ and incl/ subdirectories provide a very easy and convenient way for users to reorder table selection and output, include, or not, various table fields, and to reorder those fields in the header file by simply rearranging the various parameters as needed or desired.

\subsection{mddb, v1.3 (current)}
\label{sec:mddb}
The \ssindex{data!metadata}metadata database (mddb) package is structured in much the same way as the above described metaData package in that it is a set of classes organised around the four basic ``MIME types'' that \ssindex{projects!CyberSKA}CyberSKA currently supports in the DQS.

There are five main classes, a base class (BaseDbInsert) providing the public interface, and four classes that are sub-classed off the base class and that implement the necessary data engineering methods appropriate to the particular ``MIME type'' for which the respective classes have been developed. These relationships are shown in Table \ref{tab:mddbmods}.

\begin{table}[htbp]
  \centering
  \begin{tabular}{|l|p{8.0cm}|} 
    \hline
    \sc \textbf{Module} & \textbf{Classes} \\
    \hline
    baseDbInsert & BaseDbInsert \\
    dbFitsInsert   & DbFitsInsert(BaseDbInsert) \\
    dbUVFitsInsert& DbUVFitsInsert(BaseDbInsert)\\
    dbCImageInsert& DbCImageInsert(BaseDbInsert)\\
    dbMSInsert & DbMSInsert(BaseDbInsert)\\
    \hline
  \end{tabular}
  \caption{mddb, v1.3, Module and Class relationships}
  \label{tab:mddbmods}
\end{table}

A set of support modules provide these classes with generalised functional utility, as well as data configuration specifications, a database environment class, and setup test scripts, and shown in Table \ref{tab:mddbsubs}.

\begin{table}[htbp]
  \centering
  \begin{tabular}{|l|p{8.0cm}|} 
    \hline
    \sc \textbf{Support Directory} & \textbf{Modules} \\
    \hline
    config/ &  mddb.cfg, mddbEnv \\
    convert/ & converters, keymaps, mjdConversions \\
    db/ &  confirmKeys, xDBKeys \\
    scripts/ &  publicUris, testrun.py \\
    utils/  & f2cArgs, rUtils \\
    \hline
    docs/   &  *.html\\
    \hline
 \end{tabular}
  \caption{mddb, v1.3, Conversion, configuration, and database support}
  \label{tab:mddbsubs}
\end{table}

\section{Engineering and Workflow}
An indicated in \S \ref{sec:mddb}, data engineering is implemented by sub-classes that are specific to each dataset type. A good deal of engineering is carried out in these private methods, activities that primarily comprise evaluating and, if necessary, mitigating deprecated, errant, and/or missing \ssindex{data!metadata}metadata. Examples include converting \texttt{{\small CDELT}}\textit{i} and \texttt{{\small CROTA}}\textit{i} keywords to a \texttt{{\small CD}}\textit{i\_j} matrix as required by the \ssindex{data!model!CAOM}CAOM data model; ensuring \texttt{{\small RADESYS}} and \texttt{{\small EQUINOX}} are specified and consistent; converting a \ssindex{packages!CASA}Casa\ooindex{CASA, ascl:1107.013} Image geodetic ``telescope position'' in latitude, longitude and height to geocentric measures specified by \texttt{{\small OBSGEO-X}}, \texttt{{\small OBSGEO-Y}}, \texttt{{\small OBSGEO-Z}}. The \ssindex{projects!CyberSKA}CyberSKA project has noted that user supplied \ssindex{data!metadata}metadata can vary widely in quality and much of the data engineering so far implemented is designed to mitigate these variations to ensure compatibility with the \ssindex{data!model!CAOM}CAOM data model.

Combined in serial fashion, the metaData and mddb packages can automatically process a given astronomical dataset and make it available through\ssindex{Virtual Observatory(VO)} VO services. Since these packages process only \ssindex{data!metadata}metadata, this workflow is fast and efficient, making data\ssindex{Virtual Observatory(VO)} VO accessible in a matter of seconds, where much of that time is consumed by database interactions. Because metaData creates a textual header file, which is FITS-like in style, the mddb package does not require access to the original dataset, but only works directly from this header file. This feature has proved particularly useful in prototype development.

\section{Conclusion}
When used in conjunction, the two packages, metaData and mddb, form a workflow that is capable of receiving a dataset file in one of several formats (FITS image, UVFITS, \ssindex{packages!CASA}Casa\ooindex{CASA, ascl:1107.013} Image, Visibility Measurement Set), extract and encapsulate the dataset's \ssindex{data!metadata}metadata, write an adjudicated header file, configure those adjudicated metadata to a form appropriate to the \ssindex{data!model!CAOM}CAOM data model, and insert those data into a selected \ssindex{data!model!CAOM}CAOM database. When the database is part of a\ssindex{Virtual Observatory(VO)} VO services platform, those datasets thence become registered and published through the Canadian \ssindex{Virtual Observatory (VO)}Virtual Observatory and the larger \ssindex{Virtual Astronomical Observatory (VAO)}Virtual Astronomical Observatory network. The metaData and mddb packages have, for the first time, published \ssindex{observatories!Earth-based!ALMA}ALMA imaging data on\ssindex{Virtual Observatory(VO)} VO services. Future work aims to integrate visibility data from \ssindex{observatories!Earth-based!ALMA}ALMA and the \ssindex{observatories!Earth-based!SKA}SKA into\ssindex{Virtual Observatory(VO)} VO services and to enable user-submitted \ssindex{astronomy!radio}radio data to move seamlessly into the \ssindex{Virtual Observatory (VO)}Virtual Observatory. 

\acknowledgements This research is funded by a grant from CANARIE to the \ssindex{projects!CyberSKA}CyberSKA collaboration and a North American ALMA Science Center ALMA Development Study. This research used the facilities of the \ssindex{data centers!Canadian Astronomy Data Centre (CADC)}CADC operated by the National Research Council of Canada with the support of the Canadian Space Agency.

\bibliography{editor}
